---
title: "conda_steps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

## Create Conda Environment

```{bash}
conda create --name precip
conda activate precip
```

## List Environments

```{bash}
conda env list
```


## Create Snakemake file

```{bash}

touch Snakefile

```

### Add header variables

```
import pandas as pd
m = pd.read_csv("samples/ronca_metadata_final.csv", header = 0)
PLATES = m['plate_barcode'].unique().tolist() 
SAMPLES = m['well_barcodefull'].unique().tolist() # well barcode
```

The all rule needs to be the output you are wanting at the end...

```
rule all:
    input: 
        expand("outputs/bams/{lane}_{plate}_{sample}.sort.flt.bam", lane = LANES, plate = PLATES, sample = SAMPLES)
```

### Save one rule or sets of rules

See [this](https://github.com/lachlandeer/snakemake-econ-r/blob/master/rules/data_mgt.smk) example of one approach to saving rules or sets of rules independently, using a `.smk` extension. See this example of how to [make figures](https://github.com/lachlandeer/snakemake-econ-r/blob/master/rules/figures.smk).

In the `snakefile` we can call this as follows:

```
configfile: "paths.yaml"
runR = "Rscript --no-save --no-restore --verbose"
logAll = "2>&1"

include: config["rules"] + "data_mgt.smk"

# where a paths.yml specifies this:
rules          : "rules/"
src            : "src/"
log            : "logs/"
out            : "out/"
src_data       : "src/data/"
src_data_mgt   : "src/data-management/"
out_data       : "out/data/"
```

### Run in one `snakefile`

Otherwise, we can set up a rule as follows, specifying individual parameters and conda environments for each rule. 

 - **`benchmark`**: is a way to track a specific output within a rule to get a sense of how much memory and time it takes to run. The argument `benchmark:` is the saved output snakemake will create based on the process to create the `output` from that rule.

```
rule filter_bams:
    input: "outputs/bams/{lane}_{plate}_{sample}.sort.bam"
    output: "outputs/bams/{lane}_{plate}_{sample}.sort.flt.bam"
    conda: "envs/samtools_bwa.yml"
    threads: 1
    resources:
        mem_mb=2000,
        time=2880
    benchmark: "benchmarks/filter_bams_{lane}_{plate}_{sample}.tsv"
    shell:"""
       samtools view -f 0x2 -b {input} | samtools rmdup - {output}
       """
```

## Create DAG

We can visualize our rules or process with a DAG. The code to do this is as follows. It may require installing the `graphviz` package via brew if on mac (`brew install graphviz`) or in linux (`sudo apt-get install graphviz`).

### A filegraph

A filegraph will show what files are used and produced per each rule. A rulegraph create the graph of how rules piece together 
To create we can use:

```
snakemake --rulegraph | dot -Tpdf > rulegraph_flow.pdf
snakemake --filegraph | dot -Tpdf > filegraph_flow.pdf
snakemake --dag | dot -Tpdf > dag.pdf
```
This may require installing graphviz:
`conda install -c conda-forge graphviz`

## Run `snakemake`

We can do a dry run with:

```
snakemake --cores 1 -p -n 
```

We can run a single rule with the following. Currently this rule only works when run by itself...need to figure out order.

```
snakemake --cores 1 -p -n merge_clim_data
```

We can clean things up with:

```
snakemake --cores 1 -p -n clean_all

```

## Some Odds & Ends

There are some interesting differences between rules and how they are used. In particular, when using `shell:` vs. `script:`

- `shell:` when using this argument, we can pass additional flags to an rscript, but requires some optparse work in the Rscript itself, and must be called with `Rscript`.
- `script:` when using this argument, we can't pass additional arguments in the snakemake call...it must be a quoted path to the file (I think). But, we can access individual wildcards and params using the `snakemake@param[["csv"]]` in the R script.


### Optparse example

This comes from [here](https://github.com/lachlandeer/snakemake-econ-r/blob/master/src/data-management/gen_reg_vars.R)

```{r, eval=FALSE, echo=TRUE}

# libraries
library(optparse)
library(dplyr)

# CLI parsing
option_list = list(
    make_option(c("-d", "--data"),
                type = "character",
                default = NULL,
                help = "a csv file name",
                metavar = "character"),
    make_option(c("-p", "--param"),
                type = "numeric",
                default = 0.05,
                help = "Solow's Constant, as a numeric [default = %default]",
                metavar = "numeric"),
	make_option(c("-o", "--out"),
                type = "character",
                default = "out.csv",
                help = "output file name [default = %default]",
                metavar = "character")
);

opt_parser = OptionParser(option_list = option_list);
opt = parse_args(opt_parser);

if (is.null(opt$data)){
  print_help(opt_parser)
  stop("Input data must be provided", call. = FALSE)
}

# Load data
print("Loading data")
mrw_data <- read_csv(opt$data)
solow_const <- as.numeric(opt$param)
```

Check this rule out:

```
## rename_vars     : gives meaningful names to variables 
rule rename_vars:
    input:
        script = config["src_data_mgt"] + "rename_variables.R",
        data   = config["src_data"] + "mrw.dta",
    output:
        data = config["out_data"] + "mrw_renamed.csv",
    log:
        config["log"] + "data_cleaning/rename_variables.txt"
    shell:
        "{runR} {input.script} --data {input.data} --out {output.data} \
            > {log} {logAll}"
```

Or this to [make figures](https://github.com/lachlandeer/snakemake-econ-r/blob/master/rules/figures.smk)

```
# Rules: figures
#
# Contributors: @lachlandeer, @julianlanger, @bergmul

## make_figs: Helper rule to expand wildcard and generate all figures
rule make_figs:
    input:
        expand(config["out_figures"] + "{iFigure}.pdf",
                iFigure = PLOTS)

## figures: the recipe to make a figures using intermediate country data from MRW
rule figures:
    input:
        script = config["src_figures"] + "{iFigure}.R",
        data   = config["out_data"] + "mrw_complete.csv",
        subset = config["src_data_specs"] + "subset_intermediate.json"
    output:
        fig = config["out_figures"] + "{iFigure}.pdf",
    log:
        config["log"] + "figures/{iFigure}.txt"
    shell:
        "{runR} {input.script} \
            --data {input.data} \
            --subset {input.subset} \
            --out {output.fig} \
            > {log} {logAll}"
            
```


## Make a Dag

`snakemake --dag | dot -Tpdf > dag.pdf`

## Summary

`snakemake --summary`
